---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# yahoofinancer

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/yahoofinancer)](https://CRAN.R-project.org/package=yahoofinancer)
[![R-CMD-check](https://github.com/rsquaredacademy/yahoofinancer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rsquaredacademy/yahoofinancer/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/rsquaredacademy/yahoofinancer/branch/master/graph/badge.svg)](https://app.codecov.io/gh/rsquaredacademy/yahoofinancer?branch=master)
<!-- badges: end -->

An R interface to the Yahoo Finance API. Fetch historical prices, real-time quotes, financial statements, valuation metrics, and currency exchange rates — all returned as tidy tibbles ready for analysis.

## Features

- **Historical prices** — daily, weekly, monthly, or intraday OHLCV data
- **Real-time quotes** — current price, day range, 52-week high/low, volume
- **Financial statements** — income statement, balance sheet, and cash flow
- **Valuation metrics** — P/E, P/B, P/S, EV/EBITDA, market cap, and more
- **Currency conversion** — historical and current exchange rates
- **Market indices** — S&P 500, NIFTY 50, DAX, and any Yahoo-supported index
- **Market summary** — global snapshot of indices, commodities, and futures
- **Trending tickers** — what's popular right now, by region
- **Multi-ticker support** — batch queries via `Tickers` class or `yf_download_prices()`
- **Tidy output** — everything comes back as tibbles, ready for dplyr pipelines

## Installation

```{r gh-installation, eval = FALSE}
# Install yahoofinancer from CRAN
install.packages("yahoofinancer")

# Or the development version from GitHub
# install.packages("pak")
pak::pak("rsquaredacademy/yahoofinancer")
```

## Quick Start

```{r load_yfr, echo=FALSE, message=FALSE, warning=FALSE}
library(yahoofinancer)
```

### Download Prices

The fastest way to grab price data for one or more symbols is using the functional API:

```{r functional_api, eval=FALSE}
# Single ticker
prices <- yf_download_prices("AAPL", start = "2024-01-01", end = "2024-06-30")

# Multiple tickers at once
prices <- yf_download_prices(c("AAPL", "MSFT", "GOOGL"), period = "6mo")
```

### Single Ticker

To retrieve data from Yahoo Finance for a single stock, create an instance of the `Ticker` class by passing the company's ticker symbol as an argument:

```{r ticker}
aapl <- Ticker$new('aapl')

# get historical market data
head(aapl$get_history(start = '2024-10-20', interval = '1d'))

# meta info
# regular market price
aapl$regular_market_price

# 52 week high
aapl$fifty_two_week_high

# previous close
aapl$previous_close
```

### Multiple Tickers

You can manage multiple symbols simultaneously using the `Tickers` class. This handles errors gracefully and returns data in a long format:

```{r tickers, eval=FALSE}
tech_stocks <- Tickers$new(c("AAPL", "MSFT", "GOOGL"))

# Get current market prices for the group
tech_stocks$regular_market_price
```

### Financial Statements

Retrieve quarterly or annual financial statements using the functional API:

```{r financials, eval=FALSE}
# Get income statements
income <- yf_get_financials(c("AAPL", "MSFT"), statement_type = "income")
```

### Market Indices

To retrieve data from Yahoo Finance for an index, create an instance of the `Index` class by passing the index symbol as an argument:

```{r index}
nifty_50 <- Index$new('^NSEI')

# get historical data
head(nifty_50$get_history(start = '2024-01-20', interval = '1d'))
```

### Market Overview

Get a snapshot of the global market or see what's trending:

```{r market_overview, eval=FALSE}
# Get global market summary
market_summary <- get_market_summary()

# Get trending securities in the US
trending <- get_trending(country = "US")
```

### Currency Conversion

Retrieve current and historical exchange rates between two currencies:

```{r currency}
head(currency_converter('GBP', 'USD', '2024-01-20', '2024-01-30'))
```

## Learning More

- 📖 [Package documentation](https://yahoofinancer.rsquaredacademy.com/) — full function reference and articles
- 🐛 [Report a bug](https://github.com/rsquaredacademy/yahoofinancer/issues) — found a problem? Let us know

## Legal Disclaimer

**Yahoo!, Y!Finance, and Yahoo! finance are registered trademarks of Yahoo, Inc.**

yahoofinancer is **not** affiliated, endorsed, or vetted by Yahoo, Inc. It's an open-source tool that uses Yahoo's publicly available APIs, and is intended for research and educational purposes.

You should refer to Yahoo!'s terms of use ([here](https://policies.yahoo.com/us/en/yahoo/terms/product-atos/apiforydn/index.htm), [here](https://legal.yahoo.com/us/en/yahoo/terms/otos/index.html), and [here](https://policies.yahoo.com/us/en/yahoo/terms/index.htm)) for details on your rights to use the actual data downloaded. Remember - the Yahoo! finance API is intended for personal use only.

## Code of Conduct

Please note that the yahoofinancer project is released with a [Contributor Code of Conduct](https://yahoofinancer.rsquaredacademy.com/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
