---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# yahoofinancer

<!-- badges: start -->
[![R-CMD-check](https://github.com/rsquaredacademy/yahoofinancer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rsquaredacademy/yahoofinancer/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![Codecov test coverage](https://codecov.io/gh/rsquaredacademy/yahoofinancer/branch/master/graph/badge.svg)](https://app.codecov.io/gh/rsquaredacademy/yahoofinancer?branch=master)
<!-- badges: end -->

The goal of yahoofinancer is to obtain historical and near real time data related to stocks, index and currencies from the Yahoo Finance API.

## Installation

You can install the development version of yahoofinancer from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("rsquaredacademy/yahoofinancer")
```

## Quick Start

```{r load_yfr, echo=FALSE, message=FALSE, warning=FALSE}
library(yahoofinancer)
```


### Ticker

To retrieve data from Yahoo Finance for a single stock, create an instance of the Ticker class by passing the company's ticker symbol as an argument:

```{r}
aapl <- Ticker$new('aapl')

# get historical market data
head(aapl$get_history())

# information available in summary tab in Yahoo Finance
aapl$summary_detail
```

## IMPORTANT LEGAL DISCLAIMER 

**Yahoo!, Y!Finance, and Yahoo! finance are registered trademarks of
Yahoo, Inc.**

yahoofinancer is **not** affiliated, endorsed, or vetted by Yahoo, Inc. It's
an open-source tool that uses Yahoo's publicly available APIs, and is
intended for research and educational purposes.

**You should refer to Yahoo!'s terms of use**
([here](https://policies.yahoo.com/us/en/yahoo/terms/product-atos/apiforydn/index.htm),
[here](https://legal.yahoo.com/us/en/yahoo/terms/otos/index.html), and
[here](https://policies.yahoo.com/us/en/yahoo/terms/index.htm)) **for
details on your rights to use the actual data downloaded. Remember - the
Yahoo! finance API is intended for personal use only.**

