---
title: "Related-Symbol Recommendations and Technical Insights"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Related-Symbol Recommendations and Technical Insights}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

In this guide, you will explore the two sentiment-adjacent surfaces that Yahoo Finance exposes through `yahoofinancer`: **related-symbol recommendations** (which securities the market associates most strongly with a ticker, scored by relevance) and **technical insights** (a research snapshot combining Trading Central outlooks, key price levels, a Morningstar-style company scorecard, and an Argus Research rating with target price).

*(A note on scope: classic sell-side consensus ratings—the strong-buy/hold/sell vote distribution—are served by Yahoo's `quoteSummary` modules, which `yahoofinancer` does not currently wrap. Everything shown here comes from endpoints the package already supports.)*

### Required Packages

```{r load-libraries, message=FALSE, warning=FALSE}
# Install required packages if not already installed:
# install.packages(c("dplyr", "ggplot2", "scales", "tidyr"))

library(yahoofinancer)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(scales)
```

---

## 1. Related-Symbol Recommendations

The `recommendations` active binding on a `Ticker` object queries Yahoo's *recommendations-by-symbol* endpoint and returns a two-column frame: candidate tickers and their relevance `score` (roughly 0–1, higher means more strongly associated).

```{r fetch-recommendations}
aapl_obj <- Ticker$new("AAPL")

related <- aapl_obj$recommendations

related
#>   symbol   score
#> 1   AMZN 0.20319
#> 2   TSLA 0.19162
#> 3   GOOG 0.17865
#> 4   META 0.17095
#> 5   MSFT 0.15560
```

Scores arrive pre-sorted, but sorting explicitly makes intent clear and survives any upstream ordering change:

```{r sort-recommendations}
top_related <- related |>
  arrange(desc(score))

top_related
#>   symbol   score
#> 1   AMZN 0.20319
#> 2   TSLA 0.19162
#> ...
```

Visualize as a horizontal bar chart so long tick labels stay legible:

```{r plot-recommendations}
ggplot(top_related, aes(x = score, y = reorder(symbol, score))) +
  geom_col(fill = "#1f77b4", width = 0.65) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title    = "Securities Most Associated with AAPL",
    subtitle = "Yahoo Finance related-symbol relevance scores",
    x        = "Relevance score",
    y        = NULL,
    caption  = "Source: Yahoo Finance via yahoofinancer"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```

---

## 2. The Technical Insights Snapshot

`technical_insights` returns a nested list aggregating several research providers. It helps to pull the fields you care about into small tibbles before analyzing them. First, key technical levels from Trading Central:

```{r key-levels}
ins <- aapl_obj$technical_insights

levels_tbl <- tibble::tibble(
  level     = c("Support", "Resistance", "Stop loss"),
  price     = c(
    ins$instrumentInfo$keyTechnicals$support,
    ins$instrumentInfo$keyTechnicals$resistance,
    ins$instrumentInfo$keyTechnicals$stopLoss
  )
)

levels_tbl
#> # A tibble: 3 × 2
#>   level        price
#>   <chr>        <dbl>
#> 1 Support       417.
#> 2 Resistance    507.
#> 3 Stop loss     397.
```

Next, the short-, intermediate-, and long-term outlooks, each scored on Trading Central's evidence scale:

```{r outlooks}
te <- ins$instrumentInfo$technicalEvents

outlooks_tbl <- tibble::tibble(
  horizon     = c("Short term", "Intermediate term", "Long term"),
  direction   = c(te$shortTermOutlook$direction,
                  te$intermediateTermOutlook$direction,
                  te$longTermOutlook$direction),
  score       = c(te$shortTermOutlook$score,
                  te$intermediateTermOutlook$score,
                  te$longTermOutlook$score),
  description = c(te$shortTermOutlook$scoreDescription,
                  te$intermediateTermOutlook$scoreDescription,
                  te$longTermOutlook$scoreDescription)
)

outlooks_tbl
#> # A tibble: 3 × 4
#>   horizon           direction score description
#>   <chr>             <chr>     <dbl> <chr>
#> 1 Short term        Bullish       3 Strong Bullish Evidence
#> 2 Intermediate term Bullish       3 Strong Bullish Evidence
#> 3 Long term         Bearish       1 Weak Bearish Evidence
```

The valuation block summarizes where price sits relative to fair value:

```{r valuation}
ins$instrumentInfo$valuation$description
#> [1] "Near Fair Value"

ins$instrumentInfo$valuation$discount
#> [1] "8%"
```

---

## 3. Company Scorecard vs. Sector

The `companySnapshot` block rates the firm on six dimensions (0–1), alongside the sector median for comparison. Pivot to long form and chart both scopes side by side:

```{r company-snapshot}
snap <- ins$companySnapshot

snapshot_tbl <- tibble::tibble(
  dimension = names(snap$company),
  Company   = unlist(snap$company),
  Sector    = unlist(snap$sector)
)

snapshot_long <- snapshot_tbl |>
  pivot_longer(cols = c(Company, Sector),
               names_to = "scope",
               values_to = "score")

snapshot_long
#> # A tibble: 12 × 3
#>    dimension        scope  score
#>    <chr>            <chr>  <dbl>
#>  1 innovativeness   Company 0.977
#>  2 innovativeness   Sector  0.5
#>  3 hiring           Company 0.966
#>  4 hiring           Sector  0.5
#> # i 8 more rows
```

```{r plot-snapshot}
ggplot(snapshot_long, aes(x = score, y = reorder(dimension, score))) +
  geom_point(aes(color = scope, size = scope), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("Company" = "#1f77b4", "Sector" = "#9aa5ad")) +
  scale_size_manual(values = c("Company" = 3.4, "Sector" = 2.4), guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title    = "Company Scorecard vs. Sector Median",
    subtitle = paste("Sector:", snap$sectorInfo),
    x        = "Score",
    y        = NULL,
    color    = NULL,
    caption  = "Source: Yahoo Finance insights via yahoofinancer"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        legend.position = "top")
```

---

## 4. Research Rating and Target Price

The `recommendation` block carries a provider, categorical rating, and 12-month target price—a compact sentiment triple worth logging alongside your own models:

```{r research-call}
call_tbl <- tibble::tibble(
  provider     = ins$recommendation$provider,
  rating       = ins$recommendation$rating,
  target_price = ins$recommendation$targetPrice
)

call_tbl
#> # A tibble: 1 × 3
#>   provider       rating target_price
#>   <chr>          <chr>         <dbl>
#> 1 Argus Research BUY             620
```

Recent research report headlines are also embedded under `reports`, useful for quick qualitative context:

```{r report-headlines}
head(map_chr(ins$reports, "title"), n = 3)
#> [1] "Last week featured another monumental run for the stock market, as the major indices recaptured ..."
#> [2] "Daily - Vickers Top Buyers & Sellers for 03/10/2026"
#> ...
```

---

## 5. Comparing Related Symbols Across a Watchlist

For several symbols at once, the `Tickers` class fans out over all of them in one call. Because each source ticker contributes its own five candidates, the combined frame uses `recommended_symbol` for the candidates while `symbol` identifies the source:

```{r watchlist-fanout}
watchlist <- Tickers$new(c("AAPL", "MSFT", "GOOG"))

related_all <- watchlist$recommendations |>
  as_tibble()

related_all
#> # A tibble: 15 x 3
#>    symbol recommended_symbol   score
#>    <chr>  <chr>               <dbl>
#>  1 AAPL   AMZN              0.203
#>  2 AAPL   TSLA              0.192
#>  3 AAPL   GOOG              0.179
#>  4 MSFT   AAPL              0.212
#>  5 MSFT   ORCL              0.187
#> # i 10 more rows
```

Keep the top candidate per source ticker, sorting inside groups:

```{r watchlist-top-picks}
best_per_source <- related_all |>
  group_by(symbol) |>
  arrange(desc(score), .by_group = TRUE) |>
  slice_head(n = 1) |>
  ungroup()

best_per_source
#> # A tibble: 3 x 3
#>   symbol recommended_symbol score
#>   <chr>  <chr>              <dbl>
#> 1 AAPL   AMZN               0.203
#> 2 MSFT   AAPL               0.212
#> 3 GOOG   META               0.196
```

---

## 6. Minimal Reproducible Example

Below is the complete, self-contained workflow in a single copy-pasteable script:

```{r full-workflow}
library(yahoofinancer)
library(dplyr)
library(ggplot2)
library(scales)

# 1. Fetch and plot related-symbol recommendations
aapl_obj <- Ticker$new("AAPL")
top_related <- aapl_obj$recommendations |>
  arrange(desc(score))

ggplot(top_related, aes(x = score, y = reorder(symbol, score))) +
  geom_col(fill = "#1f77b4", width = 0.65) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title    = "Securities Most Associated with AAPL",
    subtitle = "Yahoo Finance related-symbol relevance scores",
    x        = "Relevance score",
    y        = NULL,
    caption  = "Source: Yahoo Finance via yahoofinancer"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())

# 2. Extract technical insights and key support/resistance levels
ins <- aapl_obj$technical_insights

levels_tbl <- tibble::tibble(
  level = c("Support", "Resistance", "Stop loss"),
  price = c(
    ins$instrumentInfo$keyTechnicals$support,
    ins$instrumentInfo$keyTechnicals$resistance,
    ins$instrumentInfo$keyTechnicals$stopLoss
  )
)
levels_tbl

# 3. Extract research recommendation and price target
call_tbl <- tibble::tibble(
  provider     = ins$recommendation$provider,
  rating       = ins$recommendation$rating,
  target_price = ins$recommendation$targetPrice
)
call_tbl
```

---

## 7. Summary

In this guide, you learned how to:

1. **Rank market associations**: Pull related-symbol scores with `Ticker$recommendations` and visualize them as ranked bars.
2. **Unpack research snapshots**: Flatten `technical_insights` into tidy tables of key levels, directional outlooks, and valuation posture.
3. **Compare against the sector**: Chart the six-dimension company scorecard next to its sector median.
4. **Log the research call**: Capture provider, rating, and target price as structured fields.
5. **Scale to a watchlist**: Fan out across multiple symbols with `Tickers` and rank within groups using `.by_group = TRUE`.

---

## 8. Going Further

* **Combine with prices**: Overlay `keyTechnicals` support/resistance on a `yf_download_prices()` chart—see `vignette("first-stock-analysis", package = "yahoofinancer")`.
* **Dashboard integration**: These surfaces slot directly into the reactive layer built in `vignette("shiny-dashboard", package = "yahoofinancer")`.
* **Failure semantics**: Like every network call in the package, both bindings warn and return `invisible(NULL)` when Yahoo is unreachable—guard with `is.null()` before extracting fields.
* **More recipes**: For drawdown analysis, technical indicators (EMA, RSI, MACD), and portfolio modeling, see `vignette("cookbook", package = "yahoofinancer")`.
