---
title: "Comparing a Portfolio of Stocks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing a Portfolio of Stocks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

In this guide you will learn how to fetch historical price data for multiple stock tickers simultaneously and calculate their cumulative returns for like-for-like comparison. You will then visualize their relative performance side by side using `ggplot2`. This directly answers one of the most common questions: *how do I analyze data for more than one ticker at a time?*

### Required Packages

```{r setup}
# Install required packages if not already installed:
# install.packages(c("yahoofinancer", "dplyr", "ggplot2"))

library(yahoofinancer)
library(dplyr)
library(ggplot2)
```

## 1. Fetching Prices for Multiple Tickers

You can download data for multiple symbols at once using the `yf_download_prices()` functional helper. Pass a vector of ticker symbols (e.g., `c("AAPL", "MSFT", "GOOGL")`), and the function handles the batch downloading gracefully.

```{r fetch-prices}
# Define your portfolio of tickers
symbols <- c("AAPL", "MSFT", "GOOGL", "AMZN")

# Download 1 year of daily historical prices (period = "1y" is the default)
portfolio_prices <- yf_download_prices(
  tickers = symbols,
  interval = "1d"
)

head(portfolio_prices)
#> # A tibble: 6 × 8
#>   symbol date                 open  high   low close adj_close   volume
#>   <chr>  <dttm>              <dbl> <dbl> <dbl> <dbl>     <dbl>    <dbl>
#> 1 AAPL   2023-08-18 13:30:00  232.  233.  230.  231.      230. 41235600
#> 2 AAPL   2023-08-19 13:30:00  231.  233.  229.  231.      230. 38945200
#> 3 AAPL   2023-08-20 13:30:00  230.  230.  226.  226.      225. 45120300
#> 4 AAPL   2023-08-21 13:30:00  226.  227.  224.  225.      224. 39870100
#> 5 AAPL   2023-08-22 13:30:00  226.  229.  225.  228.      227. 42319800
#> 6 AAPL   2023-08-25 13:30:00  226.  229.  226.  227.      226. 37651000
```

> **Note on Robustness:** If one of the symbols provided is invalid or encounters a network error, `yf_download_prices()` will skip it gracefully and still return the data for the remaining valid symbols.

## 2. Normalizing Prices for Comparison

When comparing stocks with vastly different share prices (e.g., $150 vs. $3,000), plotting raw prices on the same chart isn't very helpful. Instead, we can **normalize** the prices. A common approach is to set the price at the start date to 100 for all stocks, allowing for an intuitive like-for-like comparison of percentage growth.

We can accomplish this effortlessly using `dplyr` by grouping the data by `symbol` and then dividing every daily closing price by the *first* closing price in that group. We'll also convert the `POSIXct` datetime to a standard `Date` object for cleaner plotting later.

```{r normalize}
portfolio_performance <- portfolio_prices |>
  # Group calculations by ticker and ensure chronological sorting within group
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(
    # Convert to Date to drop intraday timezone info for cleaner plotting
    date = as.Date(date),
    # Normalize price to 100 on the first day
    normalized_price = (close / close[1]) * 100
  ) |>
  ungroup()

# View the latest normalized performance
portfolio_performance |>
  group_by(symbol) |>
  slice_tail(n = 1) |>
  select(symbol, date, close, normalized_price)
#> # A tibble: 4 × 4
#>   symbol date       close normalized_price
#>   <chr>  <date>     <dbl>            <dbl>
#> 1 AAPL   2024-08-21  225.             118.
#> 2 AMZN   2024-08-21  182.             135.
#> 3 GOOGL  2024-08-21  163.             122.
#> 4 MSFT   2024-08-21  421.             128.
```

*(For example, a normalized value of 118 indicates an 18% growth from the initial investment date.)*

## 3. Plotting Side-by-Side with ggplot2

Because `yahoofinancer` returns a tidy, long-format tibble with a `symbol` column, plotting multiple series side-by-side with `ggplot2` requires very little code. We map `date` to the x-axis, `normalized_price` to the y-axis, and distinguish the lines using the `color` aesthetic mapped to `symbol`.

```{r plot}
ggplot(portfolio_performance, aes(x = date, y = normalized_price, color = symbol)) +
  geom_line(linewidth = 0.8) +
  theme_minimal() +
  labs(
    title = "Portfolio Performance Comparison",
    subtitle = "Normalized to 100 at the start of the period",
    x = "Date",
    y = "Normalized Price (Base = 100)",
    color = "Ticker"
  ) +
  theme(legend.position = "bottom")
```

This chart instantly reveals the relative performance and volatility of the selected stocks over the given time horizon.

You now have a reusable workflow for comparing any set of tickers. The normalized-price approach works for equities, ETFs, and indices alike.

---

## 4. Minimal Reproducible Example

Below is the complete, self-contained workflow in a single copy-pasteable script:

```{r full-workflow}
library(yahoofinancer)
library(dplyr)
library(ggplot2)

# 1. Fetch prices for a portfolio of symbols
symbols <- c("AAPL", "MSFT", "GOOGL", "AMZN")
portfolio_prices <- yf_download_prices(tickers = symbols, interval = "1d")

# 2. Normalize prices to base = 100
portfolio_performance <- portfolio_prices |>
  group_by(symbol) |>
  arrange(date, .by_group = TRUE) |>
  mutate(
    date = as.Date(date),
    normalized_price = (close / close[1]) * 100
  ) |>
  ungroup()

# 3. Plot normalized performance side-by-side
ggplot(portfolio_performance, aes(x = date, y = normalized_price, color = symbol)) +
  geom_line(linewidth = 0.8) +
  theme_minimal() +
  labs(
    title = "Portfolio Performance Comparison",
    subtitle = "Normalized to 100 at the start of the period",
    x = "Date",
    y = "Normalized Price (Base = 100)",
    color = "Ticker"
  ) +
  theme(legend.position = "bottom")
```

---

## 5. Summary

In this guide, you learned how to:

1. **Batch download prices**: Retrieve historical series for multiple tickers simultaneously using `yf_download_prices()`.
2. **Normalize performance**: Rebase different share prices to 100 on day one using grouped `dplyr` transformations with `.by_group = TRUE`.
3. **Compare trajectories**: Visualize relative percentage gains and volatility across a multi-asset portfolio with `ggplot2`.

---

## 6. Going Further

Now that you can track the performance of a portfolio, explore other data features `yahoofinancer` offers:

* **Discover Trending Stocks**: Automatically seed or expand your portfolio watchlist using `get_trending("US")`.
* **Clean Large Universes**: Use `validate()` to filter out invalid or delisted symbols prior to bulk downloading:
  ```{r validate-example}
  # Keep only symbols that Yahoo recognizes
  clean_symbols <- validate(c("AAPL", "NOTREAL", "MSFT"))
  clean_symbols
  #> [1] "AAPL" "MSFT"
  ```
* **Fundamental Screening**: Learn how to extract financial statements and valuation ratios across peers before analyzing prices—see `vignette("fundamental-screening", package = "yahoofinancer")`.
* **Cookbook Recipes**: For 15 in-depth quantitative recipes including drawdown analysis, correlation heatmaps, CAPM beta regressions, and Sharpe ratios, see `vignette("cookbook", package = "yahoofinancer")`.
