---
title: "Currency Conversion and Forex Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Currency Conversion and Forex Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

In this guide you will learn how to discover which currencies Yahoo Finance supports and how to fetch current and historical exchange rates between any two of them. You will then combine two currency pairs (GBP/USD and EUR/USD) into a single normalized comparison chart with `ggplot2`. These workflows are unique to `yahoofinancer` — no other R package wraps Yahoo Finance's currency endpoints — and the entire guide runs in under five minutes.

### Required Packages

```{r setup}
# Install required packages if not already installed:
# install.packages(c("yahoofinancer", "dplyr", "ggplot2"))

library(yahoofinancer)
library(dplyr)
library(ggplot2)
```

## 1. Discovering Supported Currencies

Before converting anything, it helps to know what Yahoo Finance offers. The `get_currencies()` helper returns the full catalogue of supported currencies — no arguments required.

```{r get-currencies}
currencies <- get_currencies()

head(currencies)
#>   short_name         long_name symbol   local_long_name
#> 1        USD         US Dollar    USD         US Dollar
#> 2        EUR              Euro    EUR              Euro
#> 3        GBP    Pound Sterling    GBP    Pound Sterling
#> 4        JPY      Japanese Yen    JPY      Japanese Yen
#> 5        CHF       Swiss Franc    CHF       Swiss Franc
#> 6        AUD Australian Dollar    AUD Australian Dollar
```

Exchange rates on Yahoo Finance are quoted as **pairs**: to convert from one currency to another, Yahoo builds the symbol `FROM` + `TO` + `=X`. For example, British pounds to US dollars becomes `GBPUSD=X`, and euros to dollars becomes `EURUSD=X`. You rarely need to type these symbols yourself, though — the `currency_converter()` helper shown next assembles them for you.

> **Note on Robustness:** Like all `yahoofinancer` downloaders, `get_currencies()` returns `invisible(NULL)` instead of erroring when the network or API is unavailable. Assign the result and check it isn't `NULL` before continuing.

## 2. Fetching Historical Exchange Rates

`currency_converter()` retrieves current and historical rates between any two supported currencies. You can request a fixed window with `start` and `end` dates:

```{r convert-dates}
gbp_usd <- currency_converter(
  from  = "GBP",
  to    = "USD",
  start = "2024-01-01",
  end   = "2024-12-31"
)

head(gbp_usd)
#>                  date   high    low   open  close volume adj_close
#> 1 2024-01-02 00:00:00 1.2734 1.2577 1.2732 1.2602      0    1.2602
#> 2 2024-01-03 00:00:00 1.2716 1.2595 1.2678 1.2631      0    1.2631
#> 3 2024-01-04 00:00:00 1.2709 1.2602 1.2689 1.2677      0    1.2677
#> 4 2024-01-05 00:00:00 1.2721 1.2614 1.2705 1.2717      0    1.2717
#> 5 2024-01-08 00:00:00 1.2741 1.2698 1.2718 1.2720      0    1.2720
#> 6 2024-01-09 00:00:00 1.2743 1.2688 1.2694 1.2694      0    1.2694
```

The result mirrors the package's price-history schema: a data frame with `date` (POSIXct), `high`, `low`, `open`, `close`, and `volume` columns, plus `adj_close` for daily or longer intervals. A few things worth knowing:

* Instead of explicit dates, you can pass a rolling window via `period` (e.g., `period = "1mo"`); valid values range from `"1d"` to `"max"`, with `"ytd"` as the default.
* `interval` controls granularity (`"1h"`, `"1d"`, `"5d"`, `"1wk"`, `"1mo"`, `"3mo"`; default `"1d"`). Intraday intervals omit `adj_close`.
* Forex pairs trade no shares, so `volume` is reported as `0` rather than a meaningful quantity.
* `currency_converter()` is standalone — unlike most of the package, no `Ticker` object is needed.

```{r rolling-forex}
# Example: Rolling 1-month window with weekly candle intervals
weekly_rates <- currency_converter(
  from     = "GBP",
  to       = "USD",
  period   = "1mo",
  interval = "1wk"
)
```

> **Note:** For intraday analysis, remember FX markets close on weekends — hourly data will show natural gaps.

## 3. Comparing Two Pairs

The real power comes from comparing pairs side by side. Let's fetch the euro against the dollar over the same window, stack both pairs into one long data frame, and label each row with its pair.

```{r fetch-two-pairs}
eur_usd <- currency_converter(
  from  = "EUR",
  to    = "USD",
  start = "2024-01-01",
  end   = "2024-12-31"
)

fx <- bind_rows(
  gbp_usd |> mutate(pair = "GBP/USD"),
  eur_usd |> mutate(pair = "EUR/USD")
)
```

The pound trades near $1.27 while the euro trades near $1.08, so plotting raw rates would compare *levels*, not *movement*. As in the portfolio guide, we normalize each series to 100 at the start of the period so the lines show percentage change — a like-for-like view of which currency strengthened against the dollar.

```{r normalize}
fx_performance <- fx |>
  group_by(pair) |>
  arrange(date, .by_group = TRUE) |>
  mutate(
    date = as.Date(date),
    normalized_rate = (close / close[1]) * 100
  ) |>
  ungroup()

# Where did each pair end the year?
fx_performance |>
  group_by(pair) |>
  slice_tail(n = 1) |>
  select(pair, date, close, normalized_rate)
#> # A tibble: 2 × 4
#>   pair    date       close normalized_rate
#>   <chr>   <date>     <dbl>           <dbl>
#> 1 EUR/USD 2024-12-31 1.038            96.2
#> 2 GBP/USD 2024-12-31 1.252            98.6
```

*(A normalized value below 100 means the currency weakened against the dollar over the window; above 100 means it strengthened.)*

## 4. Plotting GBP/USD vs EUR/USD

Because the data is tidy — one row per date per pair — plotting both series takes a single `ggplot()` call. Map `color` to `pair` and `ggplot2` handles the rest.

```{r plot}
ggplot(fx_performance, aes(x = date, y = normalized_rate, color = pair)) +
  geom_line(linewidth = 0.8) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "grey50") +
  theme_minimal() +
  labs(
    title = "Pound vs Euro: Performance Against the US Dollar",
    subtitle = "Daily closing rates in 2024, normalized to 100 at the start of the year",
    x = "Date",
    y = "Normalized Rate (Base = 100)",
    color = "Pair"
  ) +
  theme(legend.position = "bottom")
```

In 2024 both European currencies lost ground against the dollar, but the chart makes it easy to spot divergences — months where sterling held up better than the euro, or vice versa. Swap in any other supported currencies from step 1 (e.g., `"JPY"`, `"INR"`, `"CHF"`) to rerun the exact same workflow.

You now have a reusable recipe for fetching and visualizing foreign-exchange data, from quick spot checks to multi-pair comparisons.

## Going Further

Now that you can work with exchange rates, explore related features `yahoofinancer` offers:

* The normalization technique used here is the same one covered in [Comparing a Portfolio of Stocks](comparing-portfolios.html).
* To see more practical workflows and advanced charts, browse the [yahoofinancer Cookbook](cookbook.html).
* Working with equities priced outside the US? Check the quoting currency of any ticker with the `currency` field of the `Ticker` class (e.g., `Ticker$new("RELIANCE.NS")$currency` returns `"INR"`) before mixing prices across markets.
* For full argument details and valid parameter values, see the documentation pages for `currency_converter()` and `get_currencies()`.
