---
title: "Event Study: Earnings Announcements and Price Reactions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Event Study: Earnings Announcements and Price Reactions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

In this guide, you will run a miniature **event study**: measure how a stock reacts to its own earnings announcements, after stripping out the market's general drift. The workflow is the standard one from the event-study literature, simplified to the *market-adjusted* model:

1. Pick the announcement dates (the *events*).
2. Download price history for the stock **and** a benchmark index around each date.
3. Compute daily returns; the **abnormal return** ($AR_t$) is the stock's return minus the benchmark's return on the same day.
4. Accumulate $AR_t$ across an event window (here, 2 days before to 5 days after the announcement) to obtain the **cumulative abnormal return** (CAR).

*(Note: A full event study would estimate a stock-specific beta over a pre-event estimation window and test CARs for statistical significance. The market-adjusted shortcut shown here assumes beta ≈ 1—fine for exploration and visualization, but treat any inference with care.)*

### Required Packages

```{r load-libraries, message=FALSE, warning=FALSE}
# Install required packages if not already installed:
# install.packages(c("dplyr", "tidyr", "purrr", "ggplot2", "scales"))

library(yahoofinancer)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(scales)
```

---

## 1. Define Events and Fetch Windows

We study three recent Apple earnings announcements. For each, we will download two calendar weeks on either side—enough slack that the ±2/+5 trading-day window always fits inside the fetched range:

```{r define-events}
events <- tibble::tribble(
  ~label,     ~event_date,
  "FY25 Q4",  as.Date("2025-10-30"),
  "FY26 Q1",  as.Date("2026-01-29"),
  "FY26 Q2",  as.Date("2026-04-30")
)
```

*(Note: Announcement dates are public knowledge—check Apple's investor relations page for your period of interest and adjust freely.)*

Fetching both symbols in one call keeps the benchmark perfectly aligned with the stock's calendar:

```{r fetch-single-event}
aapl_event <- yf_download_prices(
  tickers  = c("AAPL", "^GSPC"),
  start    = as.Date("2026-01-15"),
  end      = as.Date("2026-02-12"),
  interval = "1d"
)

head(aapl_event, n = 4)
#> # A tibble: 4 x 8
#>   symbol date                 open  high   low close adj_close  volume
#>   <chr>  <dttm>              <dbl> <dbl> <dbl> <dbl>     <dbl>   <dbl>
#> 1 ^GSPC  2026-01-15 14:30:00 6890. 6920. 6880. 6905.     6905. 4.90e9
#> 2 AAPL   2026-01-15 14:30:00  231.  232.  230.  231.      230. 4.51e7
#> 3 ^GSPC  2026-01-16 14:30:00 6905. 6935. 6899. 6928.     6928. 5.12e9
#> 4 AAPL   2026-01-16 14:30:00  231.  233.  230.  232.      231. 4.13e7
```

---

## 2. Compute Abnormal Returns

The helper below pivots each window to one row per day with side-by-side price columns, computes daily returns for both series, aligns every row to the announcement day (`rel_day` 0), and keeps the event window:

```{r compute-ar}
compute_ar <- function(label, event_date) {
  raw <- yf_download_prices(
    tickers  = c("AAPL", "^GSPC"),
    start    = event_date - 14,
    end      = event_date + 14,
    interval = "1d"
  )

  raw |>
    mutate(symbol = if_else(symbol == "^GSPC", "index", symbol)) |>
    select(date, symbol, close) |>
    pivot_wider(names_from = symbol, values_from = close) |>
    arrange(date) |>
    mutate(
      ret_stock = AAPL / lag(AAPL) - 1,
      ret_mkt   = index / lag(index) - 1,
      ar        = ret_stock - ret_mkt
    ) |>
    # Locate the first session on/after the announcement (t = 0)
    mutate(rel_day = row_number() - which(as.Date(date) >= event_date)[1]) |>
    filter(rel_day %in% -2:5) |>
    transmute(label, rel_day, ar)
}
```

Running it for a single event shows the shape of the result—the announcement reaction lands at `rel_day` 0 or 1 depending on whether the release came before or after the bell:

```{r inspect-ar}
ar_q1 <- compute_ar("FY26 Q1", as.Date("2026-01-29"))

ar_q1
#> # A tibble: 8 x 3
#>   label   rel_day       ar
#>   <chr>      <int>    <dbl>
#> 1 FY26 Q1       -2 -0.0042
#> 2 FY26 Q1       -1  0.0019
#> 3 FY26 Q1        0  0.0031
#> 4 FY26 Q1        1  0.0384
#> 5 FY26 Q1        2  0.0057
#> # i 3 more rows
```

---

## 3. Daily Abnormal Returns Around One Announcement

A diverging bar chart makes the announcement-day pop obvious:

```{r plot-daily-ar}
ggplot(ar_q1, aes(x = rel_day, y = ar, fill = ar >= 0)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("#c62828", "#2e7d32")) +
  scale_x_continuous(breaks = -2:5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title    = "AAPL Abnormal Returns — FY26 Q1 Earnings",
    subtitle = "Market-adjusted vs. S&P 500; day 0 = announcement",
    x        = "Trading days relative to announcement",
    y        = "Abnormal return",
    caption  = "Source: Yahoo Finance via yahoofinancer"
  ) +
  theme_minimal(base_size = 12)
```

---

## 4. Average CAR Across All Events

Now run the helper over every event with `map2_dfr()` and accumulate: thin lines show each announcement's CAR path, the thick line their average:

```{r car-across-events}
car_paths <- map2_dfr(events$label, events$event_date, compute_ar) |>
  group_by(label) |>
  arrange(rel_day, .by_group = TRUE) |>
  mutate(car = cumsum(replace_na(ar, 0))) |>
  ungroup()

avg_car <- car_paths |>
  group_by(rel_day) |>
  summarise(car = mean(car), .groups = "drop")
```

*(Note: `arrange(rel_day, .by_group = TRUE)` sorts within each event so the running sum never crosses event boundaries; `replace_na()` guards the leading `NA` return.)*

```{r plot-car}
ggplot() +
  geom_line(data = car_paths,
            aes(x = rel_day, y = car, group = label),
            color = "#9aa5ad", linewidth = 0.5) +
  geom_line(data = avg_car, aes(x = rel_day, y = car),
            color = "#1f77b4", linewidth = 1.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#555555") +
  scale_x_continuous(breaks = -2:5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    title    = "Average Cumulative Abnormal Return — AAPL Earnings",
    subtitle = "Grey: individual announcements; blue: average",
    x        = "Trading days relative to announcement",
    y        = "Cumulative abnormal return",
    caption  = "Source: Yahoo Finance via yahoofinancer"
  ) +
  theme_minimal(base_size = 12)
```

---

## 5. Reading the Results

* **A steep jump at day 0/1** means the announcement surprised the market relative to the index; a flat path implies results were priced in.
* **Drift after day 1** (continuation) versus **reversal** speaks to post-earnings-announcement behavior—compare across many quarters before believing either pattern.
* **Caveats**: single-stock CARs conflate firm news with sector news (the S&P 500 is an imperfect control); without significance testing, treat magnitudes as descriptive only. The cookbook's rolling-correlation recipe (`vignette("cookbook", package = "yahoofinancer")`) is a natural next step toward proper factor controls.

---

## 6. Minimal Reproducible Example

Below is the complete, self-contained workflow in a single copy-pasteable script:

```{r full-workflow}
library(yahoofinancer)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(scales)

# 1. Events: three earnings announcements, two weeks of slack each side
events <- tibble::tribble(
  ~label,    ~event_date,
  "FY25 Q4", as.Date("2025-10-30"),
  "FY26 Q1", as.Date("2026-01-29"),
  "FY26 Q2", as.Date("2026-04-30")
)

# 2. Market-adjusted abnormal returns over [-2, +5]
compute_ar <- function(label, event_date) {
  yf_download_prices(c("AAPL", "^GSPC"),
                     start = event_date - 14,
                     end   = event_date + 14) |>
    mutate(symbol = if_else(symbol == "^GSPC", "index", symbol)) |>
    select(date, symbol, close) |>
    pivot_wider(names_from = symbol, values_from = close) |>
    arrange(date) |>
    mutate(ret_stock = AAPL / lag(AAPL) - 1,
           ret_mkt   = index / lag(index) - 1,
           ar        = ret_stock - ret_mkt) |>
    mutate(rel_day = row_number() - which(as.Date(date) >= event_date)[1]) |>
    filter(rel_day %in% -2:5) |>
    transmute(label, rel_day, ar)
}

# 3. Cumulate per event and average across events
avg_car <- map2_dfr(events$label, events$event_date, compute_ar) |>
  group_by(label) |>
  arrange(rel_day, .by_group = TRUE) |>
  mutate(car = cumsum(replace_na(ar, 0))) |>
  ungroup() |>
  group_by(rel_day) |>
  summarise(car = mean(car), .groups = "drop")

# 4. Plot the average CAR path
ggplot(avg_car, aes(x = rel_day, y = car)) +
  geom_line(color = "#1f77b4", linewidth = 1.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#555555") +
  scale_x_continuous(breaks = -2:5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title = "Average CAR — AAPL Earnings Announcements",
       x = "Trading days relative to announcement",
       y = "Cumulative abnormal return") +
  theme_minimal(base_size = 12)
```

---

## 7. Summary

In this guide, you learned how to:

1. **Frame events**: Turn announcement dates into bounded download windows with `yf_download_prices(start, end)`.
2. **Align stock and benchmark**: Pivot both series onto a shared calendar and difference their returns.
3. **Localize reactions**: Anchor each session to the announcement via `rel_day` and slice the [-2, +5] window.
4. **Accumulate evidence**: Build per-event CAR paths (sorting with `.by_group = TRUE`) and average them into a single readable chart.

---

## 8. Going Further

* **More events**: Loop over a year of announcements—or screen other firms—and facet the CAR plots by ticker with `facet_wrap(~ symbol)`.
* **Intraday precision**: Refetch the announcement week at `interval = "60m"` (max 730-day lookback applies) to separate the opening-gap reaction from same-day drift.
* **Portfolio context**: Compare event-day volume against typical turnover using the OHLCV columns already returned.
* **More recipes**: For drawdown analysis, technical indicators (EMA, RSI, MACD, Bollinger Bands), and portfolio performance modeling, see `vignette("cookbook", package = "yahoofinancer")`.
