---
title: "Building a Live Market Dashboard with Shiny"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building a Live Market Dashboard with Shiny}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

In this guide, you will combine `yahoofinancer` with [Shiny](https://shiny.posit.co/) to build a live market dashboard: a web application that shows real-time quote cards for any ticker, its position within the 52-week range, and an interactive price history chart driven by period and interval selectors. No prior Shiny experience is required—every reactive concept used here is explained as it appears.

### Required Packages

```{r load-libraries, message=FALSE, warning=FALSE}
# Install required packages if not already installed:
# install.packages(c("yahoofinancer", "shiny", "dplyr", "ggplot2", "scales"))

library(yahoofinancer)
library(shiny)
library(dplyr)
library(ggplot2)
library(scales)
```

---

## 1. Dashboard Anatomy

Every Shiny app has two halves:

| Piece | Role | In this dashboard |
|:---|:---|:---|
| `ui` | Declares layout and widgets | Ticker picker, period/interval selects, quote cards, chart area |
| `server` | Holds the logic that reacts to inputs | Fetches quotes and price history, renders text and plots |

The bridge between them is **reactivity**: when the user changes a widget, expressions that depend on it re-execute automatically. We will use `eventReactive()` so that network calls to Yahoo Finance fire only when the user presses the *Refresh* button—polite to both the user and the API.

---

## 2. Designing the User Interface

The UI below uses `sidebarLayout()`: controls on the left, outputs on the right. Note how named vectors in `choices` display friendly labels (e.g., *"1 Month"*) while passing compact API values (`"1mo"`) to `yf_download_prices()`.

```{r ui-layout}
ui <- fluidPage(
  title = "Live Market Dashboard",
  titlePanel("Live Market Dashboard"),

  sidebarLayout(
    sidebarPanel(
      width = 3,

      selectInput(
        inputId  = "symbol",
        label    = "Ticker",
        choices  = c("AAPL", "MSFT", "GOOG", "AMZN", "NVDA", "META")
      ),

      selectInput(
        inputId = "period",
        label   = "History length",
        choices = c("1 Month" = "1mo", "3 Months" = "3mo",
                    "6 Months" = "6mo", "1 Year" = "1y",
                    "2 Years"  = "2y", "5 Years" = "5y"),
        selected = "6mo"
      ),

      selectInput(
        inputId = "interval",
        label   = "Bar interval",
        choices = c("Daily" = "1d", "Weekly" = "1wk", "Monthly" = "1mo")
      ),

      actionButton("refresh", "Refresh data", class = "btn-primary")
    ),

    mainPanel(
      width = 9,

      # Quote cards row
      fluidRow(
        column(4, wellPanel(
          h4(textOutput("price_card")),
          p(uiOutput("change_card"), style = "margin-bottom: 0;")
        )),
        column(4, wellPanel(
          h4("52-Week Range"),
          p(textOutput("range_card"), style = "margin-bottom: 0;")
        )),
        column(4, wellPanel(
          h4("Volume"),
          p(textOutput("volume_card"), style = "margin-bottom: 0;")
        ))
      ),

      # Price history chart
      plotOutput("price_chart", height = "360px")
    )
  )
)
```

---

## 3. Reactive Data Layer

The server fetches two datasets per refresh:

1. A one-row snapshot from `yf_get_market_stats()` feeding the quote cards.
2. An OHLCV series from `yf_download_prices()` feeding the chart.

Both are wrapped in `eventReactive(input$refresh, ...)`, so each press of the button triggers exactly one pair of requests.

```{r server-reactives}
server <- function(input, output, session) {

  stats <- eventReactive(input$refresh, {
    yf_get_market_stats(input$symbol)
  })

  prices <- eventReactive(input$refresh, {
    yf_download_prices(
      tickers  = input$symbol,
      period   = input$period,
      interval = input$interval
    )
  })

  # ... output renderers added in sections 4 and 5 ...
}
```

A snapshot looks like this:

```{r stats-preview}
yf_get_market_stats("AAPL")
#> # A tibble: 1 × 7
#>   symbol regular_market_price fifty_two_week_high fifty_two_week_low
#>   <chr>                 <dbl>               <dbl>              <dbl>
#> 1 AAPL                   232.                237.               169.
#> # i 3 more variables: regular_market_volume <dbl>, previous_close <dbl>,
#> #   currency <chr>
```

*(Note: Values are illustrative; live results reflect the latest available session.)*

---

## 4. Quote Cards

The cards read from the `stats()` reactive. `price_card`, `range_card`, and `volume_card` use `renderText()`, while `change_card` uses `renderUI()` to color the percentage move green or red based on direction. The range card places the current price within the 52-week band.

```{r quote-cards}
  output$price_card <- renderText({
    req(stats())
    sprintf("%s · %s", stats()$symbol, dollar(stats()$regular_market_price))
  })

  output$change_card <- renderUI({
    s <- req(stats())
    pct <- 100 * (s$regular_market_price - s$previous_close) / s$previous_close
    color <- if (pct >= 0) "#2e7d32" else "#c62828"
    tags$span(style = sprintf("color:%s", color),
              sprintf("%+.2f%% vs previous close", pct))
  })

  output$range_card <- renderText({
    s <- req(stats())
    pos <- 100 * (s$regular_market_price - s$fifty_two_week_low) /
      (s$fifty_two_week_high - s$fifty_two_week_low)
    sprintf(
      "%s — %s\n(%s of 52-week range)",
      dollar(s$fifty_two_week_low), dollar(s$fifty_two_week_high),
      percent(pos / 100, accuracy = 1)
    )
  })

  output$volume_card <- renderText({
    s <- req(stats())
    label_number(scale_cut = scales::cut_short_scale())(s$regular_market_volume)
  })
```

Two defensive details worth keeping:

* `req(stats())` silently suspends rendering until the first successful fetch—no errors flash before data arrives.
* The percentage uses `scales::percent()` with explicit `accuracy`, so values like `12.34567%` never leak into the UI.

---

## 5. Price History Chart

The chart plots closing prices over time with dashed reference lines at the 52-week extremes. Because `date` arrives as a `POSIXct` datetime, the axis works unchanged for daily, weekly, and monthly bars alike.

```{r price-chart}
  output$price_chart <- renderPlot({
    df <- req(prices())
    s  <- req(stats())

    ggplot(df, aes(x = date, y = close)) +
      geom_line(color = "#1f77b4", linewidth = 0.9) +
      geom_hline(yintercept = s$fifty_two_week_high,
                 linetype = "dashed", color = "#2e7d32") +
      geom_hline(yintercept = s$fifty_two_week_low,
                 linetype = "dashed", color = "#c62828") +
      scale_y_continuous(labels = dollar_format()) +
      labs(
        title    = paste(s$symbol, "Price History"),
        subtitle = "Dashed lines mark 52-week high and low",
        x        = NULL,
        y        = "Price (USD)",
        caption  = "Source: Yahoo Finance via yahoofinancer"
      ) +
      theme_minimal(base_size = 12) +
      theme(plot.title = element_text(face = "bold"))
  })
```

---

## 6. Minimal Reproducible Example

Here is the complete, self-contained app in a single copy-pasteable script. Save it as `app.R` and run `shiny::runApp()`.

```{r full-workflow}
library(yahoofinancer)
library(shiny)
library(dplyr)
library(ggplot2)
library(scales)

ui <- fluidPage(
  titlePanel("Live Market Dashboard"),
  sidebarLayout(
    sidebarPanel(
      width = 3,
      selectInput("symbol", "Ticker",
                  choices = c("AAPL", "MSFT", "GOOG", "AMZN", "NVDA", "META")),
      selectInput("period", "History length",
                  choices = c("1 Month" = "1mo", "3 Months" = "3mo",
                              "6 Months" = "6mo", "1 Year" = "1y",
                              "2 Years" = "2y", "5 Years" = "5y"),
                  selected = "6mo"),
      selectInput("interval", "Bar interval",
                  choices = c("Daily" = "1d", "Weekly" = "1wk",
                              "Monthly" = "1mo")),
      actionButton("refresh", "Refresh data", class = "btn-primary")
    ),
    mainPanel(
      width = 9,
      fluidRow(
        column(4, wellPanel(
          h4(textOutput("price_card")),
          p(uiOutput("change_card"), style = "margin-bottom: 0;")
        )),
        column(4, wellPanel(
          h4("52-Week Range"),
          p(textOutput("range_card"), style = "margin-bottom: 0;")
        )),
        column(4, wellPanel(
          h4("Volume"),
          p(textOutput("volume_card"), style = "margin-bottom: 0;")
        ))
      ),
      plotOutput("price_chart", height = "360px")
    )
  )
)

server <- function(input, output, session) {
  stats  <- eventReactive(input$refresh, yf_get_market_stats(input$symbol))
  prices <- eventReactive(input$refresh, {
    yf_download_prices(input$symbol, period = input$period,
                       interval = input$interval)
  })

  output$price_card <- renderText({
    s <- req(stats())
    sprintf("%s · %s", s$symbol, dollar(s$regular_market_price))
  })
  output$change_card <- renderUI({
    s <- req(stats())
    pct <- 100 * (s$regular_market_price - s$previous_close) / s$previous_close
    color <- if (pct >= 0) "#2e7d32" else "#c62828"
    tags$span(style = sprintf("color:%s", color),
              sprintf("%+.2f%% vs previous close", pct))
  })
  output$range_card <- renderText({
    s <- req(stats())
    pos <- 100 * (s$regular_market_price - s$fifty_two_week_low) /
      (s$fifty_two_week_high - s$fifty_two_week_low)
    sprintf("%s — %s\n(%s of range)",
            dollar(s$fifty_two_week_low),
            dollar(s$fifty_two_week_high),
            percent(pos / 100, accuracy = 1))
  })
  output$volume_card <- renderText({
    req(stats())$regular_market_volume |> label_number(scale_cut = cut_short_scale())
  })

  output$price_chart <- renderPlot({
    df <- req(prices())
    s  <- req(stats())
    ggplot(df, aes(x = date, y = close)) +
      geom_line(color = "#1f77b4", linewidth = 0.9) +
      geom_hline(yintercept = s$fifty_two_week_high,
                 linetype = "dashed", color = "#2e7d32") +
      geom_hline(yintercept = s$fifty_two_week_low,
                 linetype = "dashed", color = "#c62828") +
      scale_y_continuous(labels = dollar_format()) +
      labs(title = paste(s$symbol, "Price History"), x = NULL, y = "Price (USD)") +
      theme_minimal(base_size = 12) +
      theme(plot.title = element_text(face = "bold"))
  })
}

shinyApp(ui, server)
```

*(Note: Remember to press **Refresh data** after launching—the app deliberately makes no network calls until asked.)*

---

## 7. Summary

In this guide, you learned how to:

1. **Structure a Shiny app**: Pair a `fluidPage` UI with a `server` function connected by reactive IDs.
2. **Fetch on demand**: Use `eventReactive()` so Yahoo Finance is queried only when the user requests new data.
3. **Render live metrics**: Turn `yf_get_market_stats()` snapshots into formatted quote cards with `scales` labeling.
4. **Chart interactive history**: Drive `yf_download_prices()` from period/interval widgets and overlay 52-week reference levels.

---

## 8. Going Further

* **Auto-refresh**: Add `observe({ invalidateLater(60000); shinyjs::click("refresh") })` (with the `shinyjs` package) to poll every minute during market hours.
* **Multi-symbol comparison**: Fan out over several tickers at once with the [`Tickers`](../reference/Tickers.html) class—`Tickers$new(c("AAPL", "MSFT"))$get_history(period = "6mo")` returns one long tibble; sort chronologically inside groups with `arrange(date, .by_group = TRUE)` before plotting one line per `symbol`.
* **Input validation**: Guard against typos by routing custom tickers through `validate()` before fetching.
* **More recipes**: For drawdown analysis, technical indicators, and portfolio performance modeling, see `vignette("cookbook", package = "yahoofinancer")`.
